### 2023-05-21 bachelorthesis/25-gyrokinetics.tex
#39: Correction by mayself, #37: Whitspace between formula and sign

Because of the large number of particle in the fusion plasma a prediction only using Newton-Maxwell dynamics results in an impossible task for computers, but this problem can be solved with a statistical approach. For that the particle density distribution function $f(\vect{x}, \vect{v}, t)$ in the six dimensional phase space $\{\vect{x}$, $\vect{v}\}$ with the particles postion $\vect{x}$ and velocity $\vect{v}$ is needed. Because collisions are happening at much smaller frequencies than the characteristic frequenies connected to turbulence, the collisionless model is often considered \cite{Garbet2010} which gives thourgh evolution of the particle density distribution function the \textit{Vlasov equation}
\begin{gather}
	\frac{\partial f}{\partial t} + \frac{\partial f}{\partial \vect{x}} \cdot \frac{\mathrm{d}\vect{x}}{\mathrm{d}t} + \frac{\partial f}{\partial \vect{v}} \cdot \frac{\mathrm{d}\vect{v}}{\mathrm{d}t} = 0~.
	\label{eq:vlasov}
\end{gather}
To obtain a closed system the Maxwell equations with the particle density $n$ and current density $j$ can be descripted the distribution function as follows:

@@

Dynamics in tokamak plasma is descripted by spatio-temporal scales which is also referred as \textit{gyrokinetic ordering}. This seperation is a comparison between the tuburbulence frequency $\omega$ and the cyclotron frequency $\omega_\mathrm{c}$ ($\omega/\omega_\mathrm{c} \sim 10^{-3})$ due to strong external magnetic fields which causes a fast gyro motion. The length scales of the turbulence are associated with the wavevector $\vect{k}$ which can be seperated into an perpendicular component $k_\perp = |\vect{k} \times \vect{\hat{b}}|$ and a parallel component $k_\parallel = |\vect{k} \cdot \vect{\hat{b}}|$ where $\vect{\hat{b}}$ is parallel to the poloidal component of the magnetic field. The perpendicular component $k_\perp$ scales with the thermal Larmor radius $k_\perp^{-1} \sim \rhoth$ which is significant smaller as the equilibrium density $n_0$ what can be expressed with the gradient length $L_\mathrm{n} = |\nabla \ln n_0 |^{-1}$. The gradient length compares with the machnine size which leads to the normalized Larmor radius $\rho_* = \rhoth/R_0 \sim 10^{-3} - 10^{-4}$ with $R_0$ as major radius of the tokamak. The parallel component $k_\parallel$ on the other hand scales with the machine size. Experiments also show that in the core plasma the fluctuation amplitude of the density perturbation $\delta n/n_0$ and magnetic field fluctuation $\delta B/B_0$, with $B_0$ as strength of the magnetic field in equilibrium, is of order $\sim 10^{-4}$. Together this seperations result in the following result
\begin{gather}
	\frac{\omega}{\omega_\mathrm{c}} \sim \frac{k_\parallel}{k_\perp} \sim \frac{\rhoth}{L_\mathrm{n}} \sim \frac{\delta n}{n_0} \sim \frac{\delta B}{B_0} \sim \frac{v_\mathrm{d}}{\vth} \sim \epsilon_\mathrm{g}
	\label{eq:gyroordering}
\end{gather}
with $\epsilon_\mathrm{g} \ll 1$ which applies to the typical dynamics of a fusion core plasma. \cite{Brizard2007,Garbet2010} The gyrokinetic ordering allows the reduction of parameters in the Vlasov equation (\ref{eq:vlasov}).

@@

For the description for charged particle in the tokamak device the \textit{gyrocenter coordinates} will be used [Fig. \ref{fig:gyrocenterCoords}]. In this set of coordinates the gyrocenter follows the magnetic field with the parallel velocity $v_{\parallel}$. The gyro motion will be descripted together with the magnetic moment $\mu = \frac{1}{2}mv_\perp/\omega_\mathrm{c}$, the gyrocenter $\vect{X}$ and the gyro phase $\zeta$ which gives a parameter set of six quantities $\{\vect{X}, v_{\parallel}, \mu, \zeta\}$.

@@

In the next step the coordinates of the Vlasov equation (\ref{eq:vlasov}) will be transformed into the gyrocenter coordinates. For that the particle postion $\vect{x}$ will be described through the gyrocenter postion $\vect{X} = \vect{x} - \vect{\rho_\mathrm{L}}$ with the postion Larmor radius vector $\vect{\rho_\mathrm{L}}$. The particle velocity $\vect{v}$ gets decomposed into a parallel component $v_\parallel = \vect{v} \cdot \vect{\hat{b}}$ and 
a perpendicular component $v_\perp = |\vect{v} \times \vect{\hat{b}}|$ which relates to the magnetic moment $\mu$. Through the gyrokinetic ordering a phase space transformation can be performed which is based on the Lie perturbation theory, the Hamiltonian or Lagrangian formalism and an asymtotic expansion in the smallness parameter $\epsilon_\mathrm{g}$. In the lowest order is the magnetic moment an exact invariant which makes the gyro phase $\zeta$ an ignorable variable. \cite{Garbet2010,Cary1981,Cary1983} Taking everything into account the new coordinates of the Vlasov equation are $\{\vect{X}, v_{\parallel}, \mu\}$ which transferres the Vlasov equation into the \textit{gyrokinetic equation}
\begin{gather}
	\frac{\partial f}{\partial t} + \frac{\partial f}{\partial \vect{X}} \cdot \frac{\mathrm{d}\vect{X}}{\mathrm{d}t} + \frac{\partial f}{\partial v_\parallel} \cdot \frac{\mathrm{d}v_\parallel}{\mathrm{d}t} = 0~.
	\label{eq:gyrokinetic}
\end{gather}
The time derivative of the magnetic moment $\mu$ is zero because the magnetic moment is an exact invariant.

@@

All simulation in this thesis are performed with GKW (Gyro Kinetic Workshop) uses in the application of the gyrokinetic equation the $\delta f$ approximation and local limit approximation. For the $\delta f$ approximation the particle density distribution function f gets seperated into perturbation $\delta f$ and equilibrium $f_0$ (constant in time) as expressed in the gyrokinetic ordering (\ref{eq:gyroordering}). This is given by
\begin{gather}
	f = f_0 + \delta f \qquad \delta f \sim \rho_* f_0~.
\end{gather}
The distribution function in the equilibrium $f_0$ is set in GKW to the Maxwell distribution function. The perturbed distribution function $\delta f$ is considered to vary perpendicular to the poloidal magnetic field in order of the Larmor radius while the perturbation changes with the system size which results in the ordering
\begin{gather}
	\nabla_\perp(\delta f) = \nabla_\perp (f_0)~.
\end{gather}
The $\delta f$ approximation applies for global and local description where the latter case is used in this thesis.
The \textit{local limit} on the other hand uses the spatial seperation of equilibrium and perturbation and implies that all equilibrium and geometry quantities can be taken constant over the radial extent of the simulation volume. Because of the tokamak symmetry all equilibrium constants are also constant in the toridial direction where the dependency of the equilibrium magnetic field on the poloidal angle is retrained in general.

In addition to that in GKW the following coordinates get applied with the parameter $\vect{X} = \{\xcoord, \ycoord, s\}$. Here, $\xcoord$ is the radial coordinate that labels the flux surfaces normalized by the thermal Larmor radius $\rhoth$, $\ycoord$ labels the field lines and is an approximate binormal coordinate. Together with the coordinate $s$ which parameterizes the length along the field lines and is referred to as the parallel coordinate these quantities form the Hamada coordinates \cite{Hamada1958}. This results in the new coordinates $\{\xcoord, \ycoord, s, v_\parallel, \mu\}$. Combining the $\delta f$ approximation and the Hamada coordinates and the gyrokinetic equation can be formualted as follows
\begin{gather}
	\frac{\partial g}{\partial t} + \vect{v}_\chi \cdot \nabla g + (v_\parallel \vect{b} + \vect{v}_\mathrm{D}) \cdot \nabla(\delta f) - \frac{\mu B}{m}\frac{\vect{B}\cdot\nabla B}{B^2}\frac{\partial (\delta f)}{\partial v_\parallel} = S~,
\end{gather}
where $g$ is a function which contains the the perturbation $\delta f$ and the equilibrium $f_0$. Furthermore, $\vect{v}_\mathrm{D}$ and $\vect{\chi}$ are drift velocities which are caused by the inhomogeneous magnetic field and inhomogeneous magnetic field (${v}_\mathrm{D}$) and through the $\exb$ drift. The source term $S$ combines from the distribution function in the equilibrium $f_0$ and a correction term for the collisions as wall as the energy injection term. Due to prediction boundary conditions in the radial direction the radially gradients cannot change in time which is also referred as \textit{gradient driven} approach. \cite{Peeters2009}

>>>>

Because of the large number of particle in the fusion plasma a prediction only using Newton-Maxwell dynamics results in an impossible task for computers, but this problem can be solved with a statistical approach. For that the particle density distribution function $f(\vect{x}, \vect{v}, t)$ in the six dimensional phase space $\{\vect{x}$, $\vect{v}\}$ with the particles position $\vect{x}$ and velocity $\vect{v}$ is needed. Because collisions are happening at much smaller frequencies than the characteristic frequenies connected to turbulence, the collisionless model is often considered \cite{Garbet2010} which gives through evolution of the particle density distribution function the \textit{Vlasov equation}
\begin{gather}
	\frac{\partial f}{\partial t} + \frac{\partial f}{\partial \vect{x}} \cdot \frac{\mathrm{d}\vect{x}}{\mathrm{d}t} + \frac{\partial f}{\partial \vect{v}} \cdot \frac{\mathrm{d}\vect{v}}{\mathrm{d}t} = 0~.
	\label{eq:vlasov}
\end{gather}
To obtain a closed system the Maxwell equations with the particle density $n$ and current density $j$ can be descripted the distribution function as follows

@@

Dynamics in tokamak plasma is described by spatio-temporal scales which is also referred as \textit{gyrokinetic ordering}. This seperation is a comparison between the tuburbulence frequency $\omega$ and the cyclotron frequency $\omega_\mathrm{c}$ ($\omega/\omega_\mathrm{c} \sim 10^{-3})$ due to strong external magnetic fields which causes a fast gyro motion. The length scales of the turbulence are associated with the wavevector $\vect{k}$ which can be seperated into an perpendicular component $k_\perp = |\vect{k} \times \vect{b}|$ and a parallel component $k_\parallel = |\vect{k} \cdot \vect{b}|$ where $\vect{b}$ is parallel to the poloidal component of the magnetic field. The perpendicular component $k_\perp$ scales with the thermal Larmor radius $k_\perp^{-1} \sim \rhoth$ which is significant smaller as the equilibrium density $n_0$ what can be expressed with the gradient length $L_\mathrm{n} = |\nabla \ln n_0 |^{-1}$. The gradient length compares with the machine size $R$ which leads to the normalized Larmor radius $\rho_* = \rhoth/R \sim 10^{-3} - 10^{-4}$ with $R$ as major radius of the tokamak. The parallel component $k_\parallel$ on the other hand scales with the machine size $R$. Experiments also show that in the core plasma the fluctuation amplitude of the density perturbation $\delta n/n_0$ and magnetic field fluctuation $\delta B/B_0$, with $B_0$ as strength of the magnetic field in equilibrium, is of order $\sim 10^{-4}$. Together this seperations result in the following result
\begin{gather}
	\frac{\omega}{\omega_\mathrm{c}} \sim \frac{k_\parallel}{k_\perp} \sim \frac{\rhoth}{L_\mathrm{n}} \sim \frac{\delta n}{n_0} \sim \frac{\delta B}{B_0} \sim \frac{v_\mathrm{d}}{\vth} \sim \epsilon_\mathrm{g}
	\label{eq:gyroordering}
\end{gather}
with $\epsilon_\mathrm{g} \ll 1$ which applies to the typical dynamics of a fusion core plasma. \cite{Brizard2007,Garbet2010} The gyrokinetic ordering allows the reduction of parameters in the Vlasov equation (\ref{eq:vlasov}).

@@

For the description for charged particle in the tokamak device the \textit{gyrocenter coordinates} will be used [Fig. \ref{fig:gyrocenterCoords}]. In this set of coordinates the gyrocenter follows the magnetic field with the parallel velocity $v_{\parallel}$. The gyro motion will be describd together with the magnetic moment $\mu = \frac{1}{2}mv_\perp/\omega_\mathrm{c}$, the gyrocenter $\vect{X}$ and the gyro phase $\zeta$ which gives a parameter set of six quantities $\{\vect{X}, v_{\parallel}, \mu, \zeta\}$.

@@

In the next step the coordinates of the Vlasov equation (\ref{eq:vlasov}) will be transformed into the gyrocenter coordinates. For that the particle postion $\vect{x}$ will be described through the gyrocenter postion $\vect{X} = \vect{x} - \vect{\rho_\mathrm{L}}$ with the postion Larmor radius vector $\vect{\rho_\mathrm{L}}$. The particle velocity $\vect{v}$ gets decomposed into a parallel component $v_\parallel = \vect{v} \cdot \vect{b}$ and 
a perpendicular component $v_\perp = |\vect{v} \times \vect{b}|$ which relates to the magnetic moment $\mu$. Through the gyrokinetic ordering a phase space transformation can be performed which is based on the Lie perturbation theory, the Hamiltonian or Lagrangian formalism and an asymtotic expansion in the smallness parameter $\epsilon_\mathrm{g}$. In the lowest order is the magnetic moment an exact invariant which makes the gyro phase $\zeta$ an ignorable variable. \cite{Garbet2010,Cary1981,Cary1983} Taking everything into account the new coordinates of the Vlasov equation are $\{\vect{X}, v_{\parallel}, \mu\}$ which transforms the Vlasov equation into the \textit{gyrokinetic equation}
\begin{gather}
	\frac{\partial f}{\partial t} + \frac{\partial f}{\partial \vect{X}} \cdot \frac{\mathrm{d}\vect{X}}{\mathrm{d}t} + \frac{\partial f}{\partial v_\parallel} \cdot \frac{\mathrm{d}v_\parallel}{\mathrm{d}t} = 0~.
	\label{eq:gyrokinetic}
\end{gather}
The time derivative of the magnetic moment $\mu$ is zero because the magnetic moment is an exact invariant.

@@

All simulation in this thesis are performed with GKW (Gyro Kinetic Workshop) which uses in application of the gyrokinetic equation the $\delta f$ approximation and local limit approximation. For the $\delta f$ approximation the particle density distribution function f gets seperated into perturbation $\delta f$ and equilibrium $f_0$ (constant in time) as expressed in the gyrokinetic ordering (\ref{eq:gyroordering}). This is given by
\begin{gather}
	f = f_0 + \delta f \qquad \delta f \sim \rho_* f_0~.
\end{gather}
The distribution function in the equilibrium $f_0$ is set in GKW to the Maxwell distribution function. The perturbed distribution function $\delta f$ is considered to vary perpendicular to the poloidal magnetic field in order of the Larmor radius while the perturbation changes with the system size which results in the ordering
\begin{gather}
	\nabla_\perp(\delta f) = \nabla_\perp (f_0)~.
\end{gather}
The $\delta f$ approximation applies for global and local description where the latter case is used in this thesis.
The \textit{local limit} on the other hand uses the spatial seperation of equilibrium and perturbation and implies that all equilibrium and geometry quantities can be taken constant over the radial extent of the simulation volume. Because of the tokamak symmetry all equilibrium constants are also constant in the toridial direction where the dependency of the equilibrium magnetic field on the poloidal angle is retrained in general. \\\bigskip

In addition to that in GKW the following coordinates get applied with the parameter $\vect{X} = \{\xcoord, \ycoord, s\}$. Here, $\xcoord$ is the radial coordinate that labels the flux surfaces normalized by the thermal Larmor radius $\rhoth$, $\ycoord$ labels the field lines and is an approximate binormal coordinate. Together with the coordinate $s$ which parameterizes the length along the field lines and is referred to as the parallel coordinate these quantities form the Hamada coordinates \cite{Hamada1958}. This results in the new coordinates $\{\xcoord, \ycoord, s, v_\parallel, \mu\}$. Combining the $\delta f$ approximation and the Hamada coordinates the gyrokinetic equation can be formualted as follows
\begin{gather}
	\frac{\partial g}{\partial t} + \vect{v}_\chi \cdot \nabla g + (v_\parallel \vect{b} + \vect{v}_\mathrm{D}) \cdot \nabla(\delta f) - \frac{\mu B}{m}\frac{\vect{B}\cdot\nabla B}{B^2}\frac{\partial (\delta f)}{\partial v_\parallel} = S~,
\end{gather}
where $g$ is a function which contains the the perturbation $\delta f$ and the equilibrium $f_0$. Furthermore, $\vect{v}_\mathrm{D}$ and $\vect{v}_\chi$ are drift velocities which are caused by the inhomogeneous magnetic field and the perturbed magnetic field ($\vect{v}_\mathrm{D}$) and through the $\exb$ drift ($\vect{v}_\chi$). The source term $S$ combines from the distribution function in the equilibrium $f_0$ and a correction term for the collisions as well as the energy injection term. Due to prediodic boundary conditions in the radial direction the radially gradients cannot change in time which is also referred as \textit{gradient driven} approach. \cite{Peeters2009}

### 2023-05-21 bachelorthesis/32-diagnostics.tex
#39: Correction by mayself, #37: Whitspace between formula and sign

In order to diagnose the temporal evolution of the staircase pattern and to obtain an estimate of its amplitude the radial Fourier transform of the $\exb$ shearing rate is considered. 
It is defined by
\begin{equation}
	\wexb = \sum_{\kzf} \hatwexb(\kzf,t) \, \exp(\mathrm{i} \kzf \xcoord),
	\label{eq:shearingrate_fourier}
\end{equation}
where $\hatwexb$ is the complex Fourier coefficient and \linebreak $\kzf = 2\pi \nzf/L_\xcoord$
% explicitly introduce the terms zonal flow wave vector and zonal flow mode number for later usage
defines the zonal flow wave vector with the zonal flow mode number $\nzf$ ranging in $-(N_\xcoord -1)/2 \leq \nzf \leq (N_\xcoord -1)/2 $ and the radial box size $L_\xcoord$.
% define the zonal flow modes' amplitude in terms of the shearing rate
Based on the definitions above, the shear carried by the zonal flow mode with wave vector $\kzf$ is defined by $\hatwexbamp = 2 |\hatwexb(\kzf,t)|$. 
% introduces the term basic mode of the pattern for later usage
In general, the zonal flow mode that dominates the $\exb$ staircase pattern, also referred to as the \textit{basic mode} of the pattern in this work, exhibits the maximum amplitude in the spectrum $\hatwexbamp$.\\

>>>>

In order to diagnose the temporal evolution of the staircase pattern and to obtain an estimate of its amplitude the radial Fourier transform of the $\exb$ shearing rate is considered. 
It is defined by
\begin{equation}
	\wexb = \sum_{\kzf} \hatwexb(\kzf,t) \, \exp(\mathrm{i} \kzf \xcoord)~,
	\label{eq:shearingrate_fourier}
\end{equation}
where $\hatwexb$ is the complex Fourier coefficient and $\kzf = 2\pi \nzf/L_\xcoord$
% explicitly introduce the terms zonal flow wave vector and zonal flow mode number for later usage
defines the zonal flow wave vector with the zonal flow mode number $\nzf$ ranging in $-(N_\xcoord -1)/2 \leq \nzf \leq (N_\xcoord -1)/2 $ and the radial box size $L_\xcoord$.
% define the zonal flow modes' amplitude in terms of the shearing rate
Based on the definitions above, the shear carried by the zonal flow mode with wave vector $\kzf$ is defined by $\hatwexbamp = 2 |\hatwexb(\kzf,t)|$. 
% introduces the term basic mode of the pattern for later usage
In general, the zonal flow mode that dominates the $\exb$ staircase pattern, also referred to as the \textit{basic mode} of the pattern in this work, exhibits the maximum amplitude in the spectrum $\hatwexbamp$.

### 2023-05-21 bachelorthesis/34-restartscript.tex
#39: Correction by mayself, #37: Whitspace between formula and sign

As stated in Chapter \ref{sec:cluster} the \texttt{btrzx1}-cluster has a wall time of 24 hours but since simulations typical run longer than 24 hours one have to restart the simulation after the wall time gets exceeded. Some simulation need even longer so that the simulation itself needs multiple weeks to finish with a result which makes the restart of the simulation a daily task that could be easily forget. So to eliminate this minor inconveniences the following restart script were written.

As script language \texttt{python} was chosen over shell because of its easiness to write code and the variety of tools which can interact with the \texttt{bash}-shell. \texttt{bash} was also considered but after some issue occurred regarding the mechanism to check wether a file exist or not, \texttt{python} was selected.

@@

To run the restart script \texttt{python3} has to be installed on the system and should be accessible by the command line. With the following command can be the monitoring started:
\begin{lstlisting}[language=Bash]
python3 -u slurm_monitor.py --job-name $JOBNAME
\end{lstlisting}
Additional parser options can be added to the command above which will be summarized in the upcoming Chapter \ref{sub:codeFeatures}. The script was build to run in the background of the terminal for which there are two approaches:
\begin{enumerate}
    \item \texttt{nohup}: Runs command in the background while terminal can still be used or closed. \cite{nohup} The command for this method would be: \\
\begin{lstlisting}[language=Bash]
nohup python3 -u slurm_monitor.py --job-name $JOBNAME &> /dev/null &
\end{lstlisting}
and to cancel the monitoring:
\begin{lstlisting}[language=Bash]
python3 -u slurm_monitor.py --job-name $JOBNAME --kill
\end{lstlisting}
    \item \texttt{screen}: Opens a virtual terminal session which is detached from the main terminal and can be entered and closed. \cite{screen} Here the command would be:
\begin{lstlisting}[language=Bash]
screen -S $SESSION #Create screen session
python3 -u slurm_monitor.py --job-name $JOBNAME --verbose
\end{lstlisting}
and to cancel the monitoring use the keyboard shortcut \keys{\ctrl + C} or kill screen itself with \keys{\ctrl + D}. To leave the screen use ( \keys{\ctrl + A} ) + \keys{D} and to enter the screen use:
\begin{lstlisting}[language=Bash]
screen -r $SESSION
\end{lstlisting}
\end{enumerate}

@@

The restart script has several features which can be used via the command line interface as parser options. The following features with the corresponding parser option were implementing:
\begin{itemize}
    \item The script has no requirements because it runs on the standard \texttt{python3} library
    \item \textbf{\textcolor{red}{!Required!}} Set job name not longer than 8 characters to identify simulation\\
          \texttt{-j [JOBNAME], -\/-job-name [JOBNAME]}
    \item Creates jobscript file with defined content\\
          (look into variable \texttt{jobscriptContent} to add more option) \\
          \texttt{-\/-jobscript [JOBSCRIPTFILE]~~~~~~~(default=jobscript-create)}\\
          \texttt{-\/-nodes [NODES]~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~(default=3)}\\
          \texttt{-\/-ntask-per-node [TASKS]~~~~~~~~~~~~~~~~~~~~~~~~(default=32)}\\
          \texttt{-\/-walltime [WALLTIME]~~~~~~(d-hh:mm:ss)~(default=0-24:00:00)}
    \item Start/Restarts simulation until job criteria is suffused \\
          \texttt{-n [TIMESTEPS]~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\,\,(default=10000)}\\
          \texttt{-\/-restartfile [RESTARTFILE]~~~~~~~~~~~~~~~~(default=FDS.dat)}
    \item Makes backup after each run before Restart and Restore files after failed run\\
          (uses \texttt{rsync} command line utility)\\
          Quicklinks: local (simulation folder), home (home folder) \\
          \texttt{-b [LOCATION], -\/-backup [LOCATION]~~~~~~~~~~~\,(default=None)}
    \item Reset option after run fails and dump files were written\\
          (rely on \texttt{h5py}, \texttt{pandas} and \texttt{numpy} which get installed by the script itself)\\
          \texttt{-r, -\/-reset~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\,(default=False)}
    \item Creates status file with current status and progress bars and updates it dynamically\\
          \texttt{-\/-statusfile [STATUSFILE]~~~~~~~~~~~~~~~(default=status.txt)}
    \item Set form of the output table in status file\\
          Options: fancy (round box), universial (crossplattform), None (No frame)\\
          \texttt{-\/-format [FORMATTABLE]~~~~~~~~~~~~~~~~~~~~~~~~(default=None)}
    \item Sends mail at the beginning, end and restart with status file as attachment\\
          (\texttt{mailx} other equivalent has to be installed, look into \texttt{send\_mail} function for more info)\\
          \texttt{-\/-mail [MAILADDRESS]~~~~~~~~~(mail@server.de)~(default=None)}\\
          \texttt{-\/-restart-mail ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~(default=False)}
    \item Kills monitoring process by using the \texttt{PID} number\\
          \texttt{-\/-kill~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~(default=False)}
    \item Set sleep interval after which the status of the current simulation should be checked\\
          \texttt{-\/-refresh-rate [SLEEPTIME]~~~~~~~~~~~~~~~~~~~~~(default=300)}
    \item Print output of status file table to console\\
          \texttt{-v, -\/-verbose~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\,(default=False)}
\end{itemize}

>>>>

As stated in Chapter \ref{sec:cluster} the \texttt{btrzx1}-cluster has a wall time of 24 hours but since simulations typical run longer than 24 hours one have to restart the simulation after the wall time gets exceeded. Some simulation need even longer so that the simulation itself needs multiple weeks to finish which makes the restart of the simulation a daily task that could be easily forgotten. So to eliminate this minor inconveniences the following restart script were written.

As script language \texttt{python} was chosen because of its easiness to write code and the variety of tools which can interact with the \texttt{bash} shell. \texttt{bash} was also considered but after some issue occurred regarding the mechanism to check wether a file exist or not, \texttt{python} was selected. Furthermore, the script has no requirements because it runs on the standard \texttt{python3} library.

@@

To run the restart script \texttt{python3} has to be installed on the system and should be accessible via the command line. The monitoring can be started with the following command:
\begin{lstlisting}[language=Bash]
python3 -u slurm_monitor.py --job-name $JOBNAME
\end{lstlisting}
Additional parser options can be added to the command above which will be summarized in the upcoming Chapter \ref{sub:codeFeatures}. The script was build to run in the background of the terminal for which there are two approaches:
\begin{enumerate}
    \item \texttt{nohup}:\\
    Runs command in the background while terminal can still be used or closed. \cite{nohup} The command for this method would be:
\begin{lstlisting}[language=Bash]
nohup python3 -u slurm_monitor.py --job-name $JOBNAME &> /dev/null &
\end{lstlisting}
and to cancel the monitoring:
\begin{lstlisting}[language=Bash]
python3 -u slurm_monitor.py --job-name $JOBNAME --kill
\end{lstlisting}
    \item \texttt{screen}:\\
    Opens a virtual terminal session which is detached from the main terminal (like a window) and can be entered and closed. \cite{screen} Here the command would be:
\begin{lstlisting}[language=Bash]
screen -S $SESSION #Create screen session
python3 -u slurm_monitor.py --job-name $JOBNAME --verbose
\end{lstlisting}
and to cancel the monitoring use the keyboard shortcut \keys{\ctrl + C} or kill screen itself with \keys{\ctrl + D}. To leave the screen use (~\keys{\ctrl + A}~) + \keys{D} and to enter the screen use:
\begin{lstlisting}[language=Bash]
screen -r $SESSION
\end{lstlisting}
\end{enumerate}

@@

The restart script has several features which can be used via the command line interface as parser options. The following features with the corresponding parser option were implementing:
\begin{itemize}
    \item \textbf{\textcolor{red}{!Required!}} Set job name not longer than 8 characters to identify simulation\\
          \texttt{-j [JOBNAME], -\/-job-name [JOBNAME]}
    \item Creates jobscript file with defined content\\
          (look into variable \texttt{jobscriptContent} to add more option) \\
          \texttt{-\/-jobscript [JOBSCRIPTFILE]~~~~~~~(default=jobscript-create)}\\
          \texttt{-\/-nodes [NODES]~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~(default=3)}\\
          \texttt{-\/-ntask-per-node [TASKS]~~~~~~~~~~~~~~~~~~~~~~~~(default=32)}\\
          \texttt{-\/-walltime [WALLTIME]~~~~~~(d-hh:mm:ss)~(default=0-24:00:00)}
    \item Start/Restarts simulation until job criteria is suffused\\
          \texttt{-n [TIMESTEPS]~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\,\,(default=10000)}\\
          \texttt{-\/-restartfile [RESTARTFILE]~~~~~~~~~~~~~~~~(default=FDS.dat)}
    \item Makes backup after each run before Restart and Restore files after failed run\\
          (uses \texttt{rsync} command line utility)\\
          Quicklinks: local (simulation folder), home (home folder) \\
          \texttt{-b [LOCATION], -\/-backup [LOCATION]~~~~~~~~~~~\,(default=None)}
    \item Reset option after run fails and dump files were written\\
          (rely on \texttt{h5py}, \texttt{pandas} and \texttt{numpy} which get installed by the script itself)\\
          \texttt{-r, -\/-reset~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\,(default=False)}
    \item Creates status file with current status and progress bars and updates it dynamically\\
          \texttt{-\/-statusfile [STATUSFILE]~~~~~~~~~~~~~~~(default=status.txt)}
    \item Set form of the output table in status file\\
          Options: fancy (round box), universial (crossplattform), None (No frame)\\
          \texttt{-\/-format [FORMATTABLE]~~~~~~~~~~~~~~~~~~~~~~~~(default=None)}
    \item Sends mail at the beginning, end and restart with status file as attachment\\
          (\texttt{mailx} other equivalent has to be installed, look into \texttt{send\_mail} function for more info)\\
          \texttt{-\/-mail [MAILADDRESS]~~~~~~~~~(mail@server.de)~(default=None)}\\
          \texttt{-\/-restart-mail ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~(default=False)}
    \item Kills monitoring process by using the \texttt{PID} number\\
          \texttt{-\/-kill~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~(default=False)}
    \item Set sleep interval after which the status of the current simulation should be checked\\
          \texttt{-\/-refresh-rate [SLEEPTIME]~~~~~~~~~~~~~~~~~~~~~(default=300)}
    \item Print output of status file table to console\\
          \texttt{-v, -\/-verbose~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\,(default=False)}
\end{itemize}
All boolean parameters are switches, i.e., the praser option only changes the value from \texttt{False} to \texttt{True}.
